---
title: "7_income"
author: "Jen Richmond"
date: "13/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}

library(tidyverse)
library(kableExtra)
library(gt)
library(DT)

```


This week the Tidy Tuesday dataset is about income disparities in the US. I am particularly interested in the student debt dataset. 

```{r message=FALSE, warning=FALSE}
debt <- read_csv(here::here("data", "debt.csv"))

glimpse(debt)

```

Looks like year is numeric, make it a factor so that it works better in the plot 

```{r}
debt$year <- as_factor(debt$year)
```


Looks like this df average family student loan debt for aged 25-55, by race and year normalized to 2016 dollars. Lets plot debt over time by race. 

```{r}
debt %>%
  ggplot(aes(x = year, y = loan_debt, colour = race, group = race)) +
  geom_point() +
  geom_line() +
  theme_classic() +
  scale_y_continuous(limits = c(0,15000)) +
  labs(title = "Average family student load debt by race and year", 
       y = "Load debt ($2016)", 
       x = "Year")
  
```

# The goal

OK plotting is great, but I am trying to learn about tables this week. My goal is to practice making tables, lets average this data into something that is table worthy. 

This chunk groups the debt data by year (averaging across race) and summarising the meandebt levels. 
```{r}
summary <- debt %>%
  group_by(year) %>%
  summarise(meandebt = mean(loan_debt))

summary
```

I know that there are several packages that are usful for making tables in R. My goal is to try them out and work out what the pros/cons are

1. [kableExtra](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html)

2. [gt](https://gt.rstudio.com/)

3. [DT](https://rstudio.github.io/DT/)


## Option 1 kableExtra

Pipe your dataframe into the kbl() function and you get a basic html table. Add kable styling to get a bootstrap theme.

```{r}
summary %>%
  kbl()

```


```{r}
summary %>%
  kbl() %>%
  kable_styling()
```

It is weird that in each of these options, when I run the Rmd chunk it doesn't display the contents of the table. I just get a white box. But... when I knit the document, they render just fine. Not sure what is going on with that. 

I also wonder how you make it not so wide, seems to take up an excessive amount of space...


## Option 2 gt 

The gt package makes great tables (apparently). 

```{r}
summary %>%
  gt()
```

oooo that is.... minimalistic. It appears when I run the chunk and only takes up the necessary space. It is perhaps a little skinny though...

## Option 3 DT 

The DT package is an interface for the javascript DataTables library. 

```{r}
summary %>%
  DT::datatable()
```

OK that one seems nice. I don't love the default font. Its not too big (or too skinny) though and the search function could be useful if you were dealing with big tables. The decimal places are a bit of a problem though. 

# Challenges/Successes

- I got the data summary to display as a table using 3 different packages (yay!), but each one as its own problems that I will need to dig into a little more. I like the kableExtra one (except it is annoying that it won't display when I run the chunk). I like gt(), but I need to work out how to make it take up a little more room on the page. DT() is also nice, but it would be good if it would deal with those decimal places a bit more consistently. 

# Next steps

So far I have piped my data into the main table function for each of these packages and ended up with the default settings. Next I need to look at the documentaion a little more closely and work out whether there are arguments I can use to make the kableExtra table display properly, make the gt() one take up more room, and fix the decimal places and font in the DT() table. 

