tidy---
title: "Week2 transit"
date: 2021-01-05
output: html_output
---

This week I used Tidu Tuesday to play around with a custom ggplot theme. It is mostly classic with bigger fonts, more white space and shifted titles. Load it from theme_jen.R. 


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(ggeasy)
library(here)

```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2021-01-05")

```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt %>% 
  map(glimpse)

```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

cost <- tt$transit_cost


```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

cost %>%
  group_by(country) %>%
  summarise(meancost = mean(cost_km_millions)) %>%
  arrange(-meancost) %>%
  head(5) %>%
  ggplot(aes(x = reorder(country, meancost), y = meancost, fill = country)) +
  geom_col() +
  labs(y= "Average cost per km (million)", x = "Country", 
       title = "Countries with the most expensive transit projects", 
       caption = "why is the US so $$$$?") +
  theme_jen() +  # my OWN ggplot theme, mostly theme_classic with bigger fonts and more white space
  easy_remove_legend() +  # from ggeasy package
  easy_expand_y_axis()    # makes the bars sit on the x axis
                            # my OWN ggeasy-esque function, a wrapper for scale_y_continuous(expand = c(0,0))
 
   
```

```{r}

subset <- c("US", "UK", "SG", "NZ", "QA") 

cost %>%
  filter(country %in% subset) %>%
  ggplot(aes(x = length, y = cost_km_millions, colour = city)) +
  geom_point(size = 2) +
  labs(x = "Length of track (km)", y = "Cost per km (millions)", title = "Do longer routes make transit projects expensive?")


cost %>%
  filter(country %in% subset) %>%
  ggplot(aes(x = length, y = cost_km_millions, colour = city, size = tunnel)) +
  geom_point() +
  labs(x = "Length of track (km)", y = "Cost per km (millions)", title = "Do tunnels make transit projects expensive?")

cost %>%
  filter(country %in% subset) %>%
  ggplot(aes(x = length, y = cost_km_millions, colour = city, size = stations)) +
  geom_point() +
  labs(x = "Length of track (km)", y = "Cost per km (millions)", title = "Do stations make transit projects expensive?")
```

looks like just NY that is super expensive, filter out new york is the US still more expensive. 

```{r}
top5_noNY <- cost %>%
  filter(city != "New York") %>%
  group_by(country) %>%
  summarise(meancost = mean(cost_km_millions)) %>%
  arrange(-meancost) %>%
  head(5)

top5_noNY %>%
  ggplot(aes(x = reorder(country, meancost), y = meancost, fill = country)) +
  geom_col() +
  labs(y= "Cost per km (millions)", x = "Country") 
```


# custom theme

I dislike the grey/gridline style that is the default ggplot theme_grey(). theme_classic() is the closest to the APAish style I am looking for. I'd like the font to be bigger and ideally I'd like the bars to always sit on the x axis. 

```{r}
top5_noNY %>%
  ggplot(aes(x = reorder(country, meancost), y = meancost, fill = country)) +
  geom_col() +
  labs(y= "Cost per km (millions)", x = "Country") +
  theme_classic() 
```
Working on a custom theme over in theme_jen.R. Loading it here (and adding it to my RProfile)
```{r}
source(here("other_scripts", "theme_jen.R"))
```


```{r}

top5_noNY %>%
  ggplot(aes(x = reorder(country, meancost), y = meancost, fill = country)) +
  geom_col() +
  labs(y= "Cost per km (millions)", x = "Country") +
  labs(title = "Test plot", subtitle = "subtitle", caption = "caption here") +
  theme_jen() +
  easy_expand_y_axis()
    

```



```{r}
ggsave("test3.png")
```


 

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "My TidyTuesday Plot.png",
  device = "png")

```
