---
title: "Week3_artists"
author: "Jen Richmond"
date: "13/01/2021"
output: html_document
---

```{r setup, include=FALSE}


knitr::opts_chunk$set(echo = TRUE)

```

# load packages
```{r}

library(tidyverse)
library(tidytuesdayR)
library(janitor)
library(here)
library(DT)
library(feather)

source(here("other_scripts", "theme_jen.R"))

```



# Load the weekly Data

```{r}
artists <- read_csv(here("data", "w3_artist.csv"))
                    
works <- read_csv(here("data", "w3_works.csv"))
```


## Question: has the gender distribution of artists represented at the TATE changed over time?

```{r Wrangle}

artists %>%
  tabyl(gender)

```
Yikes 85% male!

What about over time, earliest artist born 1497

```{r}
artists %>%
  na.omit() %>%
  summarise(minBirth = min(year_of_birth), minDeath = min(year_of_death),
            maxBirth = max(year_of_birth), maxDeath = max(year_of_death))
```

Add new column with century

```{r}

fifteens <- 1500:1599
sixteens <- 1600:1699
seventeens <- 1700:1799
eighteens <- 1800:1899
nineteens <- 1900:1999
twenties <- 2000:2100

   artists_century <- artists %>%
  mutate(centurydeath = case_when(year_of_death %in% fifteens ~ "1500s", 
                             year_of_death %in% sixteens ~ "1600s",
                             year_of_death %in% seventeens ~ "1700s",
                             year_of_death %in% eighteens ~ "1800s",
                             year_of_death %in% nineteens ~ "1900s",
                             year_of_death %in% twenties ~ "2000s",
                             )) %>%
       mutate(centurybirth= case_when(year_of_birth %in% fifteens ~ "1500s", 
                             year_of_birth %in% sixteens ~ "1600s",
                             year_of_birth %in% seventeens ~ "1700s",
                             year_of_birth %in% eighteens ~ "1800s",
                             year_of_birth %in% nineteens ~ "1900s",
                             year_of_birth %in% twenties ~ "2000s",
                             )) 
```

```{r}
artists_century %>%
  na.omit() %>%
  tabyl(centurydeath, gender)

artists_century %>%
  tabyl(centurybirth, gender)
```

# modern artists

filter for those born since 1900

```{r}
modern <- artists %>%
  filter(year_of_birth > 1900) 

modern %>%
  tabyl(gender)
                            
```
Add decade column
```{r}
zeros <- 1900:1909
teens <- 1910:1919
two <- 1920:1929
three <- 1930:1939
four <- 1940:1949
five <- 1950:1959
six <- 1960:1969
seven <- 1970:1979
eight <- 1980:1989
nine <- 1990:1999
mill <- 2000:2009
mill_teen <- 2010:2019

 modern_decade <- modern %>%
          mutate(decade_birth = case_when(year_of_birth %in% zeros ~ "00s", 
                             year_of_birth %in% teens ~ "10s",
                             year_of_birth %in% two ~ "20s",
                             year_of_birth %in% three ~ "30s",
                             year_of_birth %in% four ~ "40s",
                             year_of_birth %in% five ~ "50s",
                              year_of_birth %in% six ~ "60s",
                              year_of_birth %in% seven ~ "70s",
                              year_of_birth %in% eight ~ "80s",
                              year_of_birth %in% nine ~ "90s",
                              year_of_birth %in% mill ~ "2000s",
                              year_of_birth %in% mill_teen ~ "2010s",
                             ))
 
 
  modern_decade$decade_birth <- as_factor(modern_decade$decade_birth)
  
 modern_decade$decade_birth <- fct_relevel(modern_decade$decade_birth, c("00s", "10s", "20s", "30s", "40s", "50s", "60s", "70s", "80s", "90s", "2000s", "2010s"))
 
```
 
 Recode NAs in gender variable. 
```{r}
 
 modern_decade$gender <- replace_na( modern_decade$gender, "Don't Know") 

modern_decade %>%
  tabyl(gender)

```




```{r define colours}

myred <-  "#cd3122" 
myblue <- "#2f7ab9"
mypink <- "#BD338F"
mygreen <- "#6c905e"

modern_decade$gender <- fct_relevel(modern_decade$gender, c("Female", "Male", "Don't Know"))

```

```{r}

modern_decade %>%
  count(decade_birth, gender) %>%
  ggplot(aes(x = decade_birth, y = n, fill = gender)) +
  geom_col() +
  scale_fill_manual(values = c(mypink, myblue, mygreen)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "Decade of Artist birth", y = "Number of artists", 
       title = "Gender distribution of the Modern artists of the Tate", subtitle = "artists born since 1900") +
 theme_jen()
  
  
```

How come there are so many NAs for gender?










# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "My TidyTuesday Plot.png",
  device = "png")

```
